### 2019-07-31 Manuscript.Rmd
Modified .Rmd file
updated .Rmd
formatted .Rmd

Most comparative studies of functional genomics have used pairwise comparisons. Yet it has been shown that this can provide biased results, since genes, like species, are phylogenetically related. Phylogenetic comparative methods should allow to correct for this, but they depend on strong assumptions, including unbiased tree estimates relative to the hypothesis being tested. An ongoing trend in comparative genomic studies is to adopt phylogenetic comparative method to answer a wide range of biological questions including evolutionary hypotheses testing. Notably among them is the recently controversial ???Ortholog Conjecture??? that assumes the functional evolution is faster in paralogs than orthologs. Using pairwise comparisons of tissue specificity index ($\tau$), earlier we provided support for the ortholog conjecture. In contrast, a recent publication suggested that the ortholog conjecture is not supported by gene expression tissue-specificity using phylogenetic independent contrasts. We find that the gene trees used suffer from important biases, due to the inclusion of trees with no duplication nodes, to the relative age of speciations and duplications, to systematic differences in branch lengths, and to non-Brownian motion of tissue-specificity on many trees. We find some support for the ortholog conjecture, but especially that incorrect implementation of phylogenetic method in empirical gene with duplications can be problematic.

@@

\vspace{0.05cm}
While all previous studies of the ortholog conjecture had used pairwise comparisons of orthologs and paralogs, a recent article suggested that this was flawed, and that phylogenetic comparative methods should be used [@RN760]. In the field of comparative biology, most of the studies still rely on the pairwise method including the testing of the ortholog conjecture. Ignorance of phylogenetic structure has been reported to underestimate the fundamental assumption of independent observations in statistics [@RN773;@RN807]. A solution is to use phylogeny-based methods to investigate such evolutionary patterns [@RN782;@RN807;@RN779;@RN780;@RN773;@RN850;@RN849]. There are three main such phylogenetic methods: Phylogenetic Independent Contrast (PIC) [@RN773], Phylogenetic Generalized Least-Square (PGLS) [@RN850], and Monte Carlo computer simulation [@RN849]. PIC is widely adopted for its relative simplicity, and its applicability to a wide range of statistical procedures [@RN781;@RN760]. The performance of PIC relies on three basic assumptions: a correct tree topology; accurate branch lengths; and trait evolution following a Brownian model (where trait variance accrues as a linear function of time)  [@RN781;@RN782;@RN783;@RN807;@RN780;@RN779;@RN816]. If any of these assumptions is incorrect, this can lead to incorrect interpretation of results. This is probably why the application of such phylogenetic methods is still limited, and debated even after being introduced about three decades ago [@RN781;@RN782;@RN783;@RN807;@RN780]. Dunn et al. [@RN760] took an innovative approach in applying PIC to compare the divergence rates between two different events ("speciation" and "duplication") to test the ortholog conjecture. However, such an application might be problematic since the time of occurrence of gene duplication, one of the two types of events compared, is unknowable by external information (e.g. no fossil evidence). Therefore, further study is required to understand why Dunn et al. [@RN760] obtained results which are inconsistent with most studies by using the phylogenetic method. It is possible that all the conclusions drawn by previous studies on gene duplication are incorrect due to overlooking phylogenetic tree structure. If so, it should be well supported.

\vspace{0.05cm}
We re-examined the data of Dunn et al., after reproducing it using the resources and scripts provided by the authors [@RN760]. Our reanalysis highlights potential problems associated with phylogenetic independent contrasts when applied to the impact of gene duplication. Finally, with proper controls, the phylogenetic method supports the ortholog conjecture.

@@

\vspace{0.5cm}
__The ortholog conjecture when $\tau$ evolution follows a Brownian model __

>>>>

Most comparative studies of functional genomics have used pairwise comparisons. Yet it has been shown that this can provide biased results, since genes, like species, are phylogenetically related. Phylogenetic comparative methods should allow to correct for this, but they depend on strong assumptions, including unbiased tree estimates relative to the hypothesis being tested. Such methods have recently been used to test the "Ortholog Conjecture", the hypothesis that functional evolution is faster in paralogs than orthologs. Whereas pairwise comparisons of tissue specificity index ($\tau$) provided support for the ortholog conjecture, it was not supported using phylogenetic independent contrasts. We find that the gene trees used suffer from important biases, due to the inclusion of trees with no duplication nodes, to the relative age of speciations and duplications, to systematic differences in branch lengths, and to non-Brownian motion of tissue-specificity on many trees. We find some support for the ortholog conjecture, but especially that incorrect implementation of phylogenetic method in empirical gene trees with duplications can be problematic.

@@

While all previous studies of the ortholog conjecture had used pairwise comparisons of orthologs and paralogs, a recent article suggested that this was flawed, and that phylogenetic comparative methods should be used [@RN760]. In the field of comparative biology, most of the studies still rely on the pairwise method including the testing of the ortholog conjecture. Ignorance of phylogenetic structure has been reported to underestimate the fundamental assumption of independent observations in statistics [@RN773;@RN807]. A solution is to use phylogeny-based methods to investigate such evolutionary patterns [@RN782;@RN807;@RN779;@RN780;@RN773;@RN850;@RN849]. There are three main such phylogenetic methods: Phylogenetic Independent Contrast (PIC) [@RN773], Phylogenetic Generalized Least-Square (PGLS) [@RN850], and Monte Carlo computer simulation [@RN849]. PIC is widely adopted for its relative simplicity, and its applicability to a wide range of statistical procedures [@RN781;@RN760]. The performance of PIC relies on three basic assumptions: a correct tree topology; accurate branch lengths; and trait evolution following a Brownian model (where trait variance accrues as a linear function of time)  [@RN781;@RN782;@RN783;@RN807;@RN780;@RN779;@RN816]. If any of these assumptions is incorrect, this can lead to incorrect interpretation of results. This is probably why the application of such phylogenetic methods is still limited, and debated even after being introduced about three decades ago [@RN781;@RN782;@RN783;@RN807;@RN780]. Dunn et al. [@RN760] took an innovative approach in applying PIC to compare the divergence rates between two different events ("speciation" and "duplication") to test the ortholog conjecture. However, such an application might be problematic since the time of occurrence of gene duplication, one of the two types of events compared, is unknowable by external information (e.g. no fossil evidence). Therefore, further study is required to understand why Dunn et al. [@RN760] obtained results which are inconsistent with most studies by using the phylogenetic method. It is possible that all the conclusions drawn by previous studies on gene duplication are incorrect due to overlooking phylogenetic tree structure. If so, it should be well supported.

We re-examined the data of Dunn et al., after reproducing it using the resources and scripts provided by the authors [@RN760]. Our reanalysis highlights potential problems associated with phylogenetic independent contrasts when applied to the impact of gene duplication. Finally, with proper controls, the phylogenetic method supports the ortholog conjecture.

@@

__The ortholog conjecture when $\tau$ evolution follows a Brownian model__
